The Spatial and Temporal Domains of Modern Ecology

### 2017-05-06 paper/SI.tex
Various updates towards revision, including new versions of figure 3

  \Hitem \emph{\textbf{What is study duration, and how do we calculate it?}} Study duration is the integral of sampling duration, or the time spent making one observation of the phenomenon in question. To clarify, duration is the total time spent sampling/observing a single point in space--not the span of time between first and last sample, nor the integral of time spent in observing all spatial replicates. 

@@

\hspace{-1cm}\underline{\textbf{Spatial scales:}}
  \Hitem \emph{\textbf{What is sampling resolution, and how do I record it?}} This is the finest scale at which a complete measurement of every unit of the quantity of interest is recorded.  For example, if the measurement in question is a tree stem count, the resolution is determined by the size of the plot used to record every tree stem. Taking this example further, let's say a study reports a plot size of 100 x 100 m, but then goes on to report that they counted stems within a single 1 m wide transect within this larger plot.  In this case, the plot resolution is in fact 100 m x 1 m, or 100 m$^2$ (sampling resolution should be reported in m$^2$).  In another example, if the reported plot size was 20 x 20 m, but the authors in fact only measured a random selection of, say, grass stems on which they counted aphids within those plots, then use an estimate of the area of the grass stem as the sampling resolution \cite{gagic_agricultural_2012}.  
  \Hitem \emph{\textbf{What is study extent, and how do I record it?}} Broadly defined, study extent is the sampling resolution multiplied by the number of spatial replicates, divided by 10,000 to convert to hectares. For studies in which the spatial replicates are not spatially contiguous (as with most field-based studies), this means plot resolution (see Q16) multiplied by number of plots. For spatially contiguous studies (e.g. those based on remote sensing imagery), it should be the total area covered by the imagery, i.e. pixel resolution multiplied by the number of pixels. However, only record the area analyzed by the authors. If they used a sub-sample of pixels, extent is the number of those pixels a multiplied by pixel resolution.

>>>>

   \Hitem \emph{\textbf{What is the effective study duration, and how do estimate it?}} The effective study duration is the total period of time over which the phenomenon of interest was observed. More specifically, in the case of repeated observations, this is the total time that elapsed between when the first and last observations at a given point were made. For once-off (unreplicated) observations, this time is equivalent to the sample duration. However, there may be cases where once-off observations may have a longer effective duration then the sample duration.  For example, consider a study that counts occurrences of pollinators over three years, using transects that are located in different locations within the broader study area during each year \cite{rollin_differences_2013-1}.  The observations are therefore not strictly temporally replicated, but the authors control for year of collection in their subsequent analysis to avoid confounding effects.  In this case, we can consider the effective duration to be three years, as the temporal information is encoded in the analysis. 
  \Hitem \emph{\textbf{What is actual study duration, and how do we calculate it?}} The actual study duration is the integral of sampling duration, or the time spent making one observation of the phenomenon in question. To clarify, duration is the total time spent sampling/observing a single point in space--not the span of time between first and last sample (the effective duration), nor the integral of time spent in observing all spatial replicates. 

@@

\hspace{-1cm}\underline{\textbf{Spatial scales:}}
  \Hitem \emph{\textbf{What is sampling resolution, and how do I record it?}} This is the finest scale at which a complete measurement of every unit of the quantity of interest is recorded.  For example, if the measurement in question is a tree stem count, the resolution is determined by the size of the plot used to record every tree stem. Taking this example further, let's say a study reports a plot size of 100 x 100 m, but then goes on to report that they counted stems within a single 1 m wide transect within this larger plot.  In this case, the plot resolution is in fact 100 m x 1 m, or 100 m$^2$ (sampling resolution should be reported in m$^2$).  In another example, if the reported plot size was 20 x 20 m, but the authors in fact only measured a random selection of, say, grass stems on which they counted aphids within those plots, then use an estimate of the area of the grass stem as the sampling resolution \cite{gagic_agricultural_2012}.  
  \Hitem \emph{\textbf{What is study extent, and how do I record it?}}  Broadly defined, actual study extent is the sampling resolution multiplied by the number of spatial replicates, divided by 10,000 to convert to hectares. For studies in which the spatial replicates are not spatially contiguous (as with most field-based studies), this means plot resolution (see Q16) multiplied by number of plots. For spatially contiguous studies (e.g., those based on remote sensing imagery), it should be the total area covered by the imagery, i.e., pixel resolution multiplied by the number of pixels. However, only record the area analyzed by the authors. If they used a sub-sample of pixels, the actual extent is the number of those pixels multiplied by pixel resolution.

Effective survey extent is defined as the total extent bounding all spatial replicates, divided by 10,000 to convert to hectares. For studies in which spatial replicates are not spatially contiguous, this means the area of the minimum convex polygon bounding all spatial replicates. For example, if the study is conducted in a national park, the effective survey extent would be the area of the national park; if the study is conducted in three national parks, the effective survey extent would be the sum of the areas of all national parks. To calculate the effective survey extent, use the area of the study area/region given in the paper; when the area is not given, but when the survey region is given by name (e.g., Joshua Tree National Park or United States), look up the area through an online search and convert to hectares. When the area is not named, but a map is given, use an appropriate digitizing platform with a suitable map-providing backend (e.g. Google Earth Pro, QGIS with OpenLayers plugin) to navigate to the region and delineate a minimum convex polygon surrounding the plots/transects/sampling units to calculate the area in hectares. When studies focus on portions of survey regions (e.g., mangroves in a coastal National Park), attempt to delineate the focal portions (the mangroves) and not the larger survey region (entire National Park) using Google Earth Pro.

For spatially contiguous studies (e.g. those based on satellite imagery), the effective survey extent is the total area covered by the imagery (in such cases, effective survey extent equals actual survey extent). However, if spatially contiguous studies only use a sub-sample of pixels, the effective survey extent is the area of the minimum convex polygon bounding all pixels (calculated following the methods above).

For studies that record individual, mobile organisms as the units of observation, use the minimum polygon surrounding the outermost observations of the complete space-time dataset (i.e. observations from all individuals and times) to define extent.

%But I'm noticing now that this isn't updated to include how to handle when the study unit is animals. I guess we could just add a sentence saying that the effective extent is the surface area of an average-sized individual of the animal species, or something along those lines. Let me know what you think.

%  Broadly defined, study extent is the sampling resolution multiplied by the number of spatial replicates, divided by 10,000 to convert to hectares. For studies in which the spatial replicates are not spatially contiguous (as with most field-based studies), this means plot resolution (see Q16) multiplied by number of plots. For spatially contiguous studies (e.g. those based on remote sensing imagery), it should be the total area covered by the imagery, i.e. pixel resolution multiplied by the number of pixels. However, only record the area analyzed by the authors. If they used a sub-sample of pixels, extent is the number of those pixels a multiplied by pixel resolution.

### 2017-05-06 paper/main.tex
Various updates towards revision, including new versions of figure 3

\usepackage[sort]{scicite}
\usepackage{times}
\usepackage{graphicx}
\usepackage{lineno}

@@

%\begin{sciabstract}
\textbf{Observational scale shapes our understanding of ecosystems \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}. Ecological patterns emerge at temporal and spatial scales that may be coarser or finer than the processes that shape them, thus investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992}. Given the centrality of scale to ecology, and the growing ability to observe ecological phenomena beyond a field perspective, it is important to understand the scales at which modern ecological research is conducted. To reveal these, we conducted a meta-analysis of recent (2004-2014) ecological research, in which we estimated the spatial resolution and extent, and temporal interval and duration, of reported ecological observations. Here we show that ecology remains a largely field-based discipline that makes observations within generally narrow spatial and temporal domains, despite the well-established literature on the importance of scale \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}. Ecologists still make limited use of remote sensing (6\% of observations), which is necessary for observing larger-scale patterns and processes \cite{turner_remote_2003,pettorelli_satellite_2014}, but make slightly greater use of technologies that capture high frequency or temporally continuous point-based measurements (12\% of observations). Resolution, extent, and duration were not clearly reported in most reviewed studies. This indicates inattentiveness to scale, which, together with the narrow spatio-temporal domains of observations, suggests a limited ability to generalize the knowledge gained in many ecological studies, which is critical for developing an ability to predict ecological responses in an era of increasing global change \cite{levin_problem_1992}. An emerging wave of observing and analytical technologies (e.g. nano-satellites, unmanned systems, open image archives, and high-power processing platforms) now makes it dramatically easier and cheaper to collect high-quality, ecologically meaningful measurements at scales inaccessible to field scientists. Combined with a greater attention to scale in research design, these advances can facilitate the spread of ecological research to sparsely observed scales, while improving the transferability of ecological knowledge.}

@@

We characterized observational domains along two key spatial dimensions, resolution (grain) and extent, and their temporal corollaries, sampling interval and total temporal duration. We calculated the actual, as opposed to representative (SI), scales for these four dimensions from 371 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 140 paper subset of 346 randomly selected articles (from 42,918 total). An additional 56 papers that were cited as the source of observations in the selected papers were also reviewed. To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI). We constructed histograms for each dimension from the mean of the perturbed ensembles, and estimated 95\% confidence intervals for each histogram bin (Fig. 1). We constructed  kernel density estimates from the full resampled ensemble in order to assess observational distributions within different juxtapositions of the four space-time dimensions (Fig. 2).

>>>>

\usepackage[sort]{scicite}
\usepackage{times}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{color}

@@

%\begin{sciabstract}
\textbf{Observational scale shapes our understanding of ecosystems \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}. Ecological patterns emerge at temporal and spatial scales that may be coarser or finer than the processes that shape them, thus investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992}. Given the centrality of scale to ecology, and the growing ability to observe ecological phenomena beyond a field perspective, it is important to understand the scales at which modern ecological research is conducted. To reveal these, we conducted a meta-analysis of recent (2004-2014) ecological research, in which we estimated the spatial resolution and extent, and temporal interval and duration, of reported ecological observations. \textcolor{blue}{Replace this text with more specific summary of observed scales:} \textcolor{red}{Here we show that ecology remains a largely field-based discipline that makes observations within generally narrow spatial and temporal domains, despite the well-established literature on the importance of scale \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}}. Ecologists still make limited use of remote sensing (6\% of observations), which is necessary for observing larger-scale patterns and processes \cite{turner_remote_2003,pettorelli_satellite_2014}, but make slightly greater use of technologies that capture high frequency or temporally continuous point-based measurements (12\% of observations). Resolution, extent, and duration were not clearly reported in most reviewed studies. This indicates inattentiveness to scale, which, together with the narrow spatio-temporal domains of observations, suggests a limited ability to generalize the knowledge gained in many ecological studies, which is critical for developing an ability to predict ecological responses in an era of increasing global change \cite{levin_problem_1992}. An emerging wave of observing and analytical technologies (e.g. nano-satellites, unmanned systems, open image archives, and high-power processing platforms) now makes it dramatically easier and cheaper to collect high-quality, ecologically meaningful measurements at scales inaccessible to field scientists. Combined with a greater attention to scale in research design, these advances can facilitate the spread of ecological research to sparsely observed scales, while improving the transferability of ecological knowledge.}

@@

We characterized observational domains along two key spatial dimensions, resolution (grain) and extent, and their temporal corollaries, sampling interval and  temporal duration (Table 1). Here spatial extent refers to the area within the perimeter described by the outer-most sample plots, and duration refers to the total time elapsed between the collection of the first and last observations a time series. We also calculated two additional measures, actual extent and actual duration, which respetively

We calculated the actual, as opposed to representative (SI), scales for these four dimensions from 375 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 134 paper subset of 346 randomly selected articles (from 42,918 total). An additional 59 papers that were cited as the source of observations in the selected papers were also reviewed.

To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI). We constructed histograms for each dimension from the mean of the perturbed ensembles, and estimated 95\% confidence intervals for each histogram bin (Fig. 1). We constructed  kernel density estimates from the full resampled ensemble in order to assess observational distributions within different juxtapositions of the four space-time dimensions (Fig. 2).
